

	A final option is, like previous deep learning successes [193, 24], to leverage the vast quantities of data available on the internet.

	. Specifically, our interest lies in internet video data. Our reasoning here is threefold:
(1) Relevant information content. In contrast to internet-scraped text or image data, internet video can uniquely
offer information regarding the physics and dynamics of the world, and information regarding human
behaviours and actions [308]. Crucially, internet video has coverage over many behaviours and tasks
relevant to a general-purpose robot (e.g., household chores).
(2) High in quantity and diversity. There are huge quantities of video data freely available on the internet [249].
Importantly, this data is highly diverse. The largest open-source robot dataset [195] pales in comparison,
both in terms of quantity and diversity of the data (see Figure 8).
(3) Internet video is relatively untapped. The use of real and simulated robot data has been extensively explored
[36, 264, 8, 127]. Meanwhile, leveraging pretrained text and image foundation models has become increasingly common in robotics [6, 157, 37, 238]. However, the use of internet video data in robotics is in more
nascent stages.

• Internet data. Robot learning has benefited from the use of broad internet data. Image and video data have
been used to pretrain visual representations for robotics [277, 188]. Foundational VLMs and LLMs have been
used to help define reward functions for the robot learner [258, 69, 317, 133]. LLMs have been employed
as high-level planners in long-horizons tasks [6, 112]. Some works have finetuned a standard internetpretrained foundation model into a robot foundation model [37, 131]. Others have jointly pretrained on
both internet data and action-labelled robot data [217, 251]. We elaborate on how internet video data has
been used to aid robot learning in Section 4.3.


- our interest lies in internet video data
    - **Relevant information content.**
        - offer information regarding the physics and dynamics of the world and information regarding human behaviours and actions [308]
        - many behaviours and tasks relevant to a general-purpose robot
    - **High in quantity and diversity.**
    - **Internet video is relatively untapped.**


	- Internet data
	    - Image and video data have been used to pretrain visual representations for robotics [277, 188].
    - Foundational VLMs and LLMs have been used to help define **reward functions** for the robot learner [258, 69, 317, 133].
    - **LLMs** have been employed as **high-level planners in long-horizons tasks** [6, 112].
    - Some works have **finetuned a standard internet-pretrained foundation model** into a robot foundation model [37, 131].
    - Others have jointly pretrained on both internet data and action-labelled robot data [217, 251].























	\item \texttt{Simulated data}
	
	Another option is to leverage simulation
	[336, 127]. However, simulation comes with issues related to flawed simulated physics and difficulties creating
	and training on a suitable diversity of simulated environments and tasks.

	- Simulated data
    - use simulations
    - Inaccuracies in low-level physics create a **‘sim-to-real’ gap** [329]
    **—> solution here is to employ domain randomization [266]
    - challenge of Manually creating a suitable diversity of simulated environments and tasks for generalist robotics 
    **—> solve this problems using procedural environment generation [62]
    - policy lack to collect high-quality data in the simulation.
    —> use of an automated curriculum during RL training [200, 158] 
    —> collecting data using specialist policies [259]
    —> models with access to privileged simulation information [92, 276, 57]
    —> bootstrapping data collection from human demonstrations [179, 177].


The use of simulation is promising for scaling up online learning while avoiding difficulties associated with the real-world [336, 127, 190]. However, the use of simulation presents a number
of issues [25]. (1) Inaccuracies in low-level physics create a ‘sim-to-real’ gap [329] that must be overcome.
A partial solution here is to employ domain randomization [266]. (2) Manually creating a suitable diversity
of simulated environments and tasks for generalist robotics is a challenge. Recent works seek to tackle this
using procedural environment generation [62, 262] and foundation-model-assisted environment design
[292, 76, 286]. (3) We often lack a policy to collect high-quality data in the simulation. Solutions here
have included the use of an automated curriculum during RL training [200, 158] and collecting data
using: specialist policies [259], models with access to privileged simulation information [92, 276, 57], or
bootstrapping data collection from human demonstrations [179, 177].

	
	
	



	% Real-world data
	One possibility is to use humans to collect robot data [36, 129]. However, this is expensive and can be difficult in tasks requiring skill.

	Data collected via human teleoperation [36, 129] and pooled data from multiple academic labs [195] have been used to train initial robot foundation models [36, 264]. Other works have investigated 	methods for automating data collection to improve scalability [35, 5, 304]. Several companies have demonstrated evidence of infrastructure suitable for large-scale robot data-collection [251, 115]. However, the largest open-source robot dataset [195] is still significantly smaller and less diverse than internet-scale data
	(see Figure 8).
	
	- Real world data
		- **Data collected via human teleoperation** [36, 129] and pooled data from multiple academic labs [195] have been used to train initial robot foundation models [36, 264].
		- methods for **automating data collection** to improve scalability [35, 5, 304].







\texttt{Low-level robot information.}
Low-level robot information refers to sensory and control signals that directly describe the robot's physical state, motion, and interaction with its environment. This includes, for instance, end-effector pose, joint positions, joint velocities, joint torques, and gripper state (e.g., open or closed). Additionally, low-level information includes information about the physical forces acting on the robot itself. In simulation environments, they may additionally include physical details of objects in the environment, such as contact forces or object dynamics.

Such low-level information is essential for learning precise motor control tasks, including grasping, locomotion, and manipulation. For example, when a robotic arm grasps a cup, it must have access to its joint angles, joint velocities, and gripper configuration in order to execute stable and accurate movements.

Similarly, while a robot may perceive a door handle through visual sensors, proprioceptive information provides the controller with knowledge about its internal state, such as how far the arm has moved, whether the wrist is correctly oriented, and whether the gripper is properly aligned with the target object.

In addition to proprioception, tactile information plays a crucial role in physical interaction. Tactile data refers to measurements obtained through direct contact with objects or surfaces and is analogous to the human sense of touch. Common tactile signals include contact forces, pressure distributions, slip detection, surface texture information, and vibration signals. These measurements are typically acquired using sensors placed on robot grippers, fingertips, or other contact surfaces.

For instance, during grasping, tactile sensors can indicate whether an object has been successfully grasped, whether it is slipping, or whether excessive force is being applied. This feedback is critical for maintaining stable and safe object manipulation.





\underline{Improvements of LfV.}

Generalization beyond 𝐷robot. LfV offers the exciting possibility of improved generalization beyond narrow
robot datasets, 𝐷robot, to the full space of tasks covered in the video dataset, 𝐷video [75]. An argument for why
this could be the case is as follows. First, consider a 𝐷robot that has good coverage over the low-level skills or
‘atomic’ actions required from the robot (e.g., specific grasping motions or locomotion skills) [53]. Now consider
a task unseen in 𝐷robot but seen in 𝐷video. Our combined LfV dataset (𝐷LfV = {𝐷robot, 𝐷video}) may contain most
information required to complete the task. 𝐷robot provides information regarding how to execute the low-level
robot skills, whilst the human actions in 𝐷video provides higher-level information regarding how to complete the
overall task (e.g., visual information regarding required movements and steps). Thus, a suitable LfV method may
be capable of leveraging 𝐷video to generalise beyond 𝐷robot and solve the task. There is preliminary evidence of
such generalization in LfV-related literature [37, 67, 289, 278]. Figure 2 explores this generalization setting in
more detail



Emergent Capabilities. Learning from internet-scale video may yield capabilities qualitatively beyond what can
be obtained when learning only from a narrow 𝐷robot. We expect this for two reasons. First, in other domains large
quantities of internet data have allowed for unexpected ‘emergent’ capabilities [212, 39]. Second, diverse internet
video paired with language annotations offers a path towards stitching together the lower-level knowledge
obtained from robot and video data with the rich abstractions and world knowledge that can be obtained from
text data [193].
Improvements in-distribution of 𝐷robot. Finally, we also expect video data to yield improvements in tasks that
are in-distribution of the robot dataset. Utilising a large video dataset can allow for improved data-efficiency
with respect to 𝐷robot [188]. Additionally, LfV approaches may obtain higher absolute task performance (e.g.,
higher success rates) in settings in-distribution of 𝐷robot, versus non-LfV approaches [289].







\texttt{Challenges of Video Modalities.}
Although internet-scale video data offers significant potential for robotic learning, it also introduces several fundamental challenges. In contrast to robot interaction datasets, videos are typically high-dimensional, noisy, weakly labeled, and highly diverse. Extracting meaningful representations and task-relevant information from such data is therefore a non-trivial problem.

\texttt{Missing Action Labels.}
One major limitation of raw video data is the absence of explicit action labels required by many imitation learning and offline reinforcement learning approaches. While videos provide rich visual observations of behaviors, they typically do not contain the low-level control commands executed during the task. Retrospectively inferring robot actions from human videos is challenging, since human motions often cannot be directly mapped to the action space of a robotic system. Furthermore, videos usually lack additional reinforcement learning signals such as reward annotations, goal labels, or start-of- and end-of-episode boundaries.



Another challenge arises from the distribution shift between internet videos and the downstream robotics domain. Differences may occur in camera viewpoints, environments, embodiments, object appearances, and task distributions. In particular, internet videos predominantly contain human demonstrations, while the downstream agent is often a robot with different kinematics, morphology, and control capabilities. This embodiment gap makes it difficult to directly transfer observed human motions to robot actions. Human demonstrations may additionally contain behaviors that are inefficient, ambiguous, or irrelevant for robotic execution. Furthermore, some important robotic tasks, such as specialized industrial manipulation tasks, may be underrepresented in publicly available video datasets.



\texttt{Distribution Shift.}
Another challenge arises from the distribution shift between internet videos and the downstream robotics domain. Differences may occur in camera viewpoints, environments, embodiments, object appearances, and task distributions. The Human demonstrations in videos may contain behaviors that are inefficient, ambiguous, or irrelevant for robotic execution. Auch muss die embodiments gap bezüglich der Ausführungsfrom zwischen Mensch und roboter berücksichtigt werden, da der roboter different kinematics, morphology, and control capabilities hat. Und es muss die visuelle Lücke zwischen den Task-Operater im Video und den Roboterarm geschlossen werden. The embodiment gap makes it difficult to directly transfer observed human motions to robot actions
Additionally, some important robotic tasks, such as specialized industrial manipulation tasks, may be underrepresented in publicly available video datasets. These discrepancies make transferring knowledge from human videos to robotic systems particularly difficult.



Controllability, Stochasticity, and Partial Observability. In unlabelled video, it can be difficult to disentangle
which parts of a transition are affected (i.e., controlled) by a specific agent’s actions or which are due to the external
environment or noise [104]. This can be problematic for methods that attempt to extract action information from
video (Section 4.2.1). Furthermore, the stochastic nature and partial-















Internet videos contain large amounts of semantic, physical, and behavioral information that can be useful for robotics. However, these videos are typically unlabeled and differ significantly from robot observations, making knowledge extraction challenging. Recent work addresses this problem using video foundation models (FMs), which are large deep learning models trained on large-scale internet video datasets. These models learn general video representations that can be transferred to robotics tasks through finetuning or adaptation. In addition, techniques developed for video FMs can be combined with robot data to improve representation learning and generalization. Current approaches mainly focus on video encoding, video prediction, and video-to-text generation methods.


```latex id="w7h3pk"
\texttt{Extracting Foundational Knowledge from Internet Video.}
A central challenge in Learning from Video (LfV) is how to extract robotics-relevant knowledge from large-scale and heterogeneous internet video data. Unlike robot datasets, internet videos typically lack action labels and contain significant distribution shifts with respect to the downstream robotics domain. Nevertheless, such videos contain rich information about object interactions, physical behaviors, task structures, and human activities that can be valuable for robotic learning.

Recent progress in \textbf{Video Foundation Models (Video FMs)} provides a promising direction for addressing this challenge. Video foundation models are large-scale deep learning models trained on massive internet video datasets. Through large-scale training, these models learn general visual and temporal representations that capture semantics, motion dynamics, physical interactions, and behavioral patterns. As a result, they can extract foundational knowledge that is transferable across a wide range of downstream tasks, including robotics applications.

\texttt{Using pretrained FMs:} One important application of Video FMs is the transfer of pretrained models to robotics tasks. For example, a video prediction model trained on internet video data may be adapted or fine-tuned into a robot dynamics model, allowing the robot to leverage the broad gathered knowledge about object motion and environmental dynamics from the pretrained Video FM. 

\texttt{Using techniques and datasets inspired by Video FMs:} In addition, techniques developed for video foundation models can be integrated into robot learning pipelines and can lead to better extraction of representations and knowledge. This, in turn, can lead to improved generalization and task understanding. For instance, a Robot FM can be directly trained by combining robot data with large-scale video inspired by video FMs techniques [251].


The success of foundation models in natural language processing and computer vision further motivates this direction. Similar to how text and image foundation models have improved reasoning, representation learning, and generalization in robotics, video foundation models may provide robots with richer knowledge about temporal dynamics, human behavior, and physical interactions. Consequently, extracting foundational knowledge from internet video is considered a key component for developing scalable and generalizable Learning from Video systems.
```











\paragraph{Overview of Video Representation Learning}

Video representation learning aims to learn a video encoder 
$p(z \mid X)$ that maps a video clip $X$ to a latent representation $z$. 
The learned representation should capture semantic and temporal information 
useful for downstream tasks such as robotic control and imitation learning.

Let
\begin{itemize}
    \item $X$ denote a video clip,
    \item $\bar X$ denote a video clip paired with additional information 
    such as labels, object masks, bounding boxes, language annotations, or metadata,
    \item $Y$ denote a learning target derived from $\bar X$.
\end{itemize}

During training, the encoder learns representations that enable the prediction 
or alignment of $Y$ from the input video $X$.

\paragraph{Prediction-based Objectives}

Prediction-based approaches learn representations by predicting a target $Y$ 
from the input video $X$. Depending on the task, $Y$ may correspond to future 
video frames, masked observations, motion information, or other supervision signals.

For example, in video prediction tasks, the target is defined as the next frame

\[
Y = X_{t+1},
\]

which is predicted from previous observations $X_{\leq t}$.

The encoder is optimized using a prediction objective of the form

\[
\mathcal{L}_{\text{pred}}
=
\| \hat{Y} - Y \|^2,
\]

where $\hat{Y}$ denotes the predicted target.

\paragraph{Joint-Embedding Objectives}

Joint-embedding approaches first map both the video input $X$ and the target $Y$ 
into a shared latent embedding space:

\[
z_X = f(X),
\qquad
z_Y = g(Y).
\]

A loss function is then applied to align semantically similar embeddings:

\[
\mathcal{L}_{\text{align}}
=
L(z_X, z_Y).
\]

Contrastive learning is one of the most widely used approaches in this setting. 
A common contrastive objective is the InfoNCE loss:

\[
\mathcal{L}_{\text{NCE}}
=
-
\log
\frac{
\exp(\mathrm{sim}(z_X, z_Y)/\tau)
}{
\sum_k
\exp(\mathrm{sim}(z_X, z_k)/\tau)
}.
\]























\subsection{Video Prediction Models}


\paragraph{Methods.}
Modern video prediction approaches are dominated by diffusion models, autoregressive transformers, and masked transformers.

\textbf{Diffusion models} generate videos via iterative denoising and are effective at modeling complex, high-dimensional continuous distributions. However, they suffer from slow sampling and limited long-horizon consistency. Efficiency is improved by operating in latent spaces rather than pixel space, and by leveraging large-scale image pretraining. Recent methods also explore direct full-sequence generation to improve temporal coherence.


\textbf{Autoregressive and masked transformers} operate in discrete latent spaces, typically obtained via vector-quantized autoencoders (VQ-AE). Autoregressive models factorize video generation sequentially, which leads to exposure bias and error accumulation over long horizons. Masked transformers instead predict missing tokens in parallel using iterative refinement (e.g., MaskGIT-style decoding), offering improved efficiency and reduced drift, making them attractive for long-horizon robotics tasks.

\textbf{Conditioning mechanisms} are critical for controllability. VPMs can be conditioned on actions, language, or goal images, enabling structured control over generated futures:
\[
p(\tau \mid c).
\]
Action conditioning enables counterfactual simulation of control strategies, language conditioning supports high-level task specification, and image conditioning enables goal-directed prediction. In robotics, such conditioning effectively turns video prediction models into controllable world models, bridging perception, planning, and imitation learning.







```latex
\subsection{Video Prediction Models}

Video prediction models (VPMs) aim to learn a predictive distribution over future visual observations conditioned on past context. The standard formulation is next-frame prediction:
\[
p(o_{t+1} \mid o_{t-k:t}),
\]
while more generally VPMs can be expressed as conditional video generation models:
\[
p(\tau \mid c),
\]
where $\tau$ denotes a video sequence and $c$ represents conditioning information such as past observations, actions, language instructions, or goal images. When trained on large-scale internet video data, these models learn rich priors over physical dynamics, object interactions, and human behavior. As a result, they provide a useful foundation for downstream robotic learning tasks, particularly imitation learning and model-based control.

\subsubsection{Methods}

\paragraph{General Overview of Video Prediction.}
Video prediction models aim to learn a mapping from past observations (and optionally actions) to future video sequences. Let
\begin{itemize}
    \item $X = o_{t-k:t}$ denote observed video frames,
    \item $A = a_{t-k:t}$ denote optional action inputs,
    \item $Y = o_{t+1:t+T}$ denote future frames to be predicted.
\end{itemize}

The learning objective is to model:
\[
p(Y \mid X, A).
\]

During training, the model learns to predict future visual outcomes from past context, either in pixel space or in a learned latent space:
\[
(X, A) \rightarrow Y.
\]

\paragraph{Diffusion Models.}
Diffusion-based video prediction models generate future frames via iterative denoising. They are particularly effective at modeling complex continuous distributions and producing high-fidelity samples. However, they suffer from two main limitations: slow sampling speed and difficulty maintaining long-horizon temporal consistency.

To improve efficiency, many approaches perform diffusion in a learned latent space rather than pixel space. This reduces computational cost while preserving semantic structure. In addition, training often leverages large-scale image datasets alongside video data to improve visual quality and generalization.

\paragraph{Autoregressive and Masked Transformer Models.}
Transformer-based approaches typically operate in a discrete latent space learned via vector-quantized autoencoders (VQ-AE). This provides a tokenized representation of video that is suitable for sequence modeling.

Autoregressive models factorize the joint distribution as:
\[
p(x) = \prod_t p(x_t \mid x_{<t}),
\]
and generate tokens sequentially. While expressive, these models suffer from exposure bias, where errors accumulate over long rollouts due to conditioning on previous predictions.

Masked transformer models instead predict missing tokens in parallel and iteratively refine the output (e.g., MaskGIT-style decoding). Compared to autoregressive approaches, masked models are more computationally efficient and reduce temporal drift, making them more suitable for long-horizon video prediction in robotics.

\paragraph{Conditioning Mechanisms.}
Conditioning plays a central role in simplifying video prediction and enabling controllability, which is crucial for robotic applications. In general, VPMs model:
\[
p(\tau \mid c),
\]
where different conditioning signals define different control interfaces.

Common conditioning modalities include action conditioning, language conditioning, and goal image conditioning. Action conditioning enables simulation of alternative control strategies, language conditioning provides a flexible interface for task specification, and goal conditioning enables direct prediction of task-aligned future outcomes. 

From a robotics perspective, conditioning effectively turns video prediction models into controllable world models, bridging perception, planning, and imitation learning.
```





















In diffusion-based video prediction, future video frames are generated by iteratively denoising a noisy representation of the target video sequence. During training, Gaussian noise is progressively added to real future observations, and the model learns to reverse this corruption process conditioned on context information such as previous frames, actions, or language instructions.

In pixel-space diffusion, noise is added directly to raw video frames. In latent diffusion models, the video frames are first encoded into a lower-dimensional latent representation, and and diffusion is performed in this compressed latent space instead of directly on pixels. 

The denoised latent representation is then decoded back into video frames. Operating in latent space reduces computational cost and removes low-level pixel redundancy while preserving the semantic and temporal structure of the video.


During inference, diffusion-based video prediction models start from a randomly sampled noisy representation and iteratively denoise it into a plausible future video sequence. The denoising process is conditioned on contextual information such as previous observations, actions, or language instructions. In latent diffusion models, this iterative generation process is performed in a compressed latent space rather than directly in pixel space, after which a decoder reconstructs the final video frames.




\paragraph{Diffusion Models.}
Diffusion-based video prediction models generate future frames via iterative denoising a latent representation or the raw pixel of a video. They are particularly effective at modeling complex continuous distributions and producing high-fidelity samples. However, they suffer from two main limitations: slow sampling speed and difficulty maintaining long-horizon temporal consistency.

To improve efficiency, many approaches perform diffusion in a learned latent space rather than pixel space. This reduces computational cost while preserving semantic structure. In addition, training often leverages large-scale image datasets alongside video data to improve visual quality and generalization.






\paragraph{Autoregressive and Masked Transformer Models.}
Autoregressive and masked transformer models provide an alternative approach to video prediction compared to diffusion models. Rather than generating future frames through iterative denoising, transformer-based approaches model video as a sequence of discrete tokens, typically obtained from a vector-quantized autoencoder (VQ-AE). 

Autoregressive transformers generate future tokens sequentially by predicting the next token conditioned on previously generated tokens. While effective at modeling temporal dependencies, prediction errors can accumulate over long horizons, leading to temporal drift. In contrast, masked transformer models predict multiple masked tokens in parallel and iteratively refine the generated sequence, improving computational efficiency and reducing long-horizon error accumulation.

Compared to diffusion models, transformer-based approaches generally require fewer inference steps and can therefore achieve faster generation. However, diffusion models often produce higher visual fidelity and better capture complex continuous distributions.


\paragraph{Autoregressive and Masked Transformer Models.}
Rather than generating future frames through iterative denoising like diffusion models, transformer-based approaches model videos as a sequence of discrete tokens, typically obtained from a vector-quantized autoencoder (VQ-AE).

Autoregressive transformers generate future tokens sequentially by predicting the next token conditioned on previously generated tokens. While effective at modeling temporal dependencies, small prediction errors can accumulate over long horizons, leading to temporal drift and reduced generation quality. 
Masked transformer models instead predict multiple masked tokens in parallel and iteratively refine the generated sequence, improving computational efficiency and reducing error accumulation compared to autoregressive decoding.

Compared to diffusion models, transformer-based approaches generally require fewer inference steps and can therefore achieve faster generation. Diffusion models, however, often produce higher visual fidelity and better model complex continuous distributions, although they can also struggle to maintain temporal consistency during long-horizon video generation.



Transformer-based video prediction models typically operate on discrete visual tokens rather than directly on raw pixels. To obtain these tokens, the video frames are first compressed using a Vector-Quantized Autoencoder (VQ-AE). The encoder maps the input video into a lower-dimensional latent representation, which is then quantized by assigning each latent feature to the nearest entry in a learned codebook of discrete embeddings. As a result, the video can be represented as a sequence of discrete visual tokens, analogous to words in natural language processing.

Autoregressive transformer models generate future video tokens sequentially by predicting the next token conditioned on all previously generated tokens. This allows the model to capture temporal dependencies in video sequences. However, because each prediction depends on previous generated outputs, small prediction errors can accumulate over long horizons, leading to temporal drift and degraded generation quality.

Masked transformer models address this limitation by masking subsets of video tokens and predicting multiple missing tokens in parallel instead of generating them strictly one-by-one. The model iteratively refines the masked predictions until a complete video sequence is reconstructed. Compared to autoregressive decoding, masked transformers improve computational efficiency and reduce long-horizon error accumulation.


Transformer-based video prediction models typically operate on discrete visual tokens rather than directly on raw pixels. To obtain these tokens, the video frames are first compressed using a Vector-Quantized Autoencoder (VQ-AE). The encoder maps the input video into a lower-dimensional continuous latent representation. These latent features are then quantized by replacing each feature vector with the nearest embedding vector from a learned codebook. The selected codebook indices form a discrete token representation of the video, analogous to word tokens in natural language processing.

Autoregressive transformer models generate future video tokens sequentially by predicting the next token conditioned on all previously generated tokens.
 This allows the model to capture temporal dependencies in video sequences. However, because each prediction depends on previous generated outputs, small prediction errors can accumulate over long horizons, leading to temporal drift and degraded generation quality.

Masked transformer models address this limitation by masking subsets of video tokens and predicting multiple missing tokens in parallel instead of generating them strictly one-by-one. The model iteratively refines the masked predictions until a complete video sequence is reconstructed. 
During inference, the model can be conditioned on a short history of recent camera observations from the robot environment, which are encoded into visible tokens, while the tokens corresponding to future time steps are initially masked. The model then predicts the missing future tokens for the entire horizon in parallel and iteratively refines these predictions until a coherent future video sequence is obtained.
The model can also be conditioned on additional inputs such as language instructions, goal images, or robot state information, allowing it to generate future video predictions that are aligned with a desired task or behavior.
Compared to autoregressive decoding, masked transformers improve computational efficiency and reduce long-horizon error accumulation.




















---

#### High-quality representations
Video-to-text models learn rich spatio-temporal representations by aligning visual sequences with language. Compared to video-only models, they tend to develop more semantically structured latent representations. Compared to image-to-text models, they explicitly capture temporal dynamics, such as motion patterns, causality, and action progression.

These representations are particularly valuable for robotics, where understanding *state transitions over time* is essential for control. In practice, robotics systems can leverage these representations through:
- representation transfer from pretrained video-language models
- multimodal pretraining that incorporates robot interaction data into large-scale video-language corpora

This aligns with the broader trend of using foundation models as perceptual backbones for downstream robotic policies.

---

#### Grounded reasoning and planning
Large language models (LLMs) have demonstrated strong capabilities as high-level planners in robotics. However, their lack of direct grounding in physical perception limits their reliability in real-world environments.

Video-to-text models address this limitation by conditioning reasoning directly on rich visual-temporal input. This enables:
- improved grounding of symbolic reasoning in observed physical states
- better modeling of object interactions and environment dynamics
- more consistent closed-loop reasoning, where predictions are continuously informed by perceptual feedback

As a result, they can serve as intermediate modules between perception and action, supporting more reliable long-horizon planning compared to text-only LLM-based planners.

---

#### Annotating robot data
Large-scale robotic datasets often require costly manual annotation, especially for semantic labels such as actions, goals, or task descriptions.

Video-to-text models can be used as automated annotation systems to generate:
- step-by-step natural language descriptions of demonstrations
- action labels or sub-task segmentation
- goal-conditioned summaries of trajectories

This is particularly relevant for imitation learning, where high-quality language annotations can improve:
- behavior cloning performance
- dataset interpretability
- data filtering and curriculum design

Such capabilities are increasingly used to scale robotic datasets without proportional human labeling effort.

---

#### Rewards and preference signals
Beyond annotation, video-to-text models can also be used to derive reward signals or value estimates for reinforcement learning and imitation learning systems.

This can be achieved through:
- visual question answering formulations (e.g., “Did the robot successfully grasp the object?”)
- learned evaluators that score trajectories based on semantic criteria
- RL from AI feedback (RLAIF), where the model acts as a surrogate reward function

In this setting, video-to-text models act as *learned critics*, providing dense or high-level feedback that is otherwise difficult to specify manually. This is particularly useful in sparse-reward robotic environments.

---

#### Limitations and considerations
Despite their advantages, video-to-text models also introduce several challenges in robotic applications:

- **Hallucination risk:** generated language may not faithfully reflect physical reality
- **Temporal ambiguity:** fine-grained action boundaries are often difficult to extract reliably
- **Computational cost:** large video-language models are expensive to train and run in real time
- **Distribution shift:** performance may degrade significantly when moving from internet video data to embodied robot domains

These limitations motivate tighter integration with physical interaction data and robotics-specific fine-tuning strategies.

---

Overall, video-to-text models provide a powerful interface between perception, language, and decision-making, making them a key component in modern imitation learning and robot foundation model pipelines.






















To address these challenges, alternative action representations $\hat{a} \in \hat{A}$ can be derived from video data. These representations serve as proxy supervision signals that enable learning from action-free videos and facilitate video-to-robot transfer under missing action labels and domain shift, as illustrated in Figure~\ref{fig:RecoveringActionFepresentationsFromVideo_1}.

\paragraph{Alternative Action Representations $\hat{a}$}

In many real-world video datasets, explicit robot action labels (e.g., joint velocities or torques) are not available. Recovering such low-level actions directly from video is an ill-posed problem due to ambiguity and the lack of physical interaction information.

To overcome this limitation, recent approaches introduce an alternative action representation space $\hat{A}$, which defines a latent or abstract action domain. Instead of modeling raw robot control signals, actions are represented in a semantic feature space that captures high-level behavior such as grasping, moving, or pushing.

Formally, an alternative action representation is defined as
\[
\hat{a} \in \hat{A},
\]
where $\hat{A}$ is a learned or predefined latent action space derived from visual or multimodal video data. The elements $\hat{a}$ do not correspond directly to robot motor commands, but encode behavior-relevant semantics that are easier to infer from video.

These representations act as an intermediate bridge between visual observations and robot control, enabling policies to be learned in $\hat{A}$ and later mapped to the robot's true action space.


















\paragraph{Challenges in Video-to-Robot Transfer}

The main challenges in transferring knowledge from video data to robotic systems are:
\begin{itemize}
    \item (i) \textbf{Missing action labels}: Most video datasets do not contain explicit robot action annotations, which are required for learning control policies and dynamics models.
    \item (ii) \textbf{Distribution shift}: There exists a significant domain gap between human or internet videos and robot execution. This includes differences in visual appearance, embodiment, dynamics, and behavior patterns, which can severely degrade the generalization of models trained on video data when applied to robotic tasks.
\end{itemize}

\paragraph{Problem Formulation}

Formally, video data provides only state transitions of the form $(o_t, o_{t+1})$, without access to the corresponding actions $a_t$. This makes it difficult to directly apply standard imitation learning or reinforcement learning methods, which typically require triplets $(o_t, a_t, o_{t+1})$.

\paragraph{Alternative Action Representations}

To address this limitation, recent approaches propose to infer \textbf{alternative action representations} $\hat{a} \in \hat{A}$ from video data. These representations serve as proxy supervision signals that encode the underlying behavior generating a transition.

Instead of attempting to recover true robot control signals directly from video—which is inherently ill-posed due to ambiguity and missing physical interaction information—the idea is to define a latent action space $\hat{A}$ that captures higher-level, semantically meaningful behavior such as \textit{grasping}, \textit{pushing}, or \textit{moving an object}.

Formally, this induces a relabeling of video transitions as:
\[
(o_t, o_{t+1}) \;\rightarrow\; (o_t, \hat{a}_t, o_{t+1}), \quad \hat{a}_t \in \hat{A}.
\]

Here, $\hat{A}$ can be interpreted as a learned or predefined latent action space derived from visual or multimodal representations.

\paragraph{Learning in the Alternative Action Space}

The key idea is that $\hat{a}$ acts as a surrogate for missing action labels, enabling learning from action-free video data. A policy can then be defined in this alternative space as:
\[
\pi_{\hat{A}}(\hat{a}_t \mid o_t),
\]
which maps observations to latent action representations rather than low-level motor commands.

To execute such actions on a physical robot, a mapping function is required that translates latent actions into the robot’s actual action space:
\[
a_t = f_{\text{align}}(\hat{a}_t),
\]
where $f_{\text{align}}$ bridges the gap between the learned latent representation and real robot control signals.

\paragraph{Intuition}

Intuitively, this framework introduces an intermediate level of abstraction between perception and control. A video encoder extracts visual information about \textit{what is happening}, while the latent action space $\hat{A}$ represents \textit{what is being done} in a semantic sense, independent of embodiment-specific motor details.

\begin{itemize}
    \item The \textbf{video encoder} extracts visual state representations.
    \item The \textbf{alternative action space $\hat{A}$} encodes semantic behaviors such as \textit{reach}, \textit{grasp}, or \textit{lift}.
    \item The \textbf{policy $\pi_{\hat{A}}$} predicts latent actions from observations.
    \item The \textbf{alignment function $f_{\text{align}}$} maps latent actions to executable robot commands.
\end{itemize}

This decomposition enables learning from large-scale, unlabelled video data while mitigating both the missing-action problem and the domain shift between human videos and robotic systems.